---
title: "Homework 5"
author: "Shreyasi Tamhane, eid: syt333, GitHub Repo: "
date: "2024-02-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, results = 'hide', echo = FALSE, warning=FALSE, message=FALSE}
# import libraries
library(ggplot2)
library(tidyverse)
library(mosaic)
library(kableExtra)
# import datasets
letter_freq <- read.csv("letter_frequencies.csv")
```

### Problem 1: Iron Bank

```{r, echo=FALSE, results = 'hide'}
illegal_trade_sim = do(100000)*nflip(n=2021, prob=0.024)
sum(illegal_trade_sim >= 70) # p-val: 0.0023
210/100000
```

**Null Hypothesis**

The null hypothesis that we are testing is that securities trades from the Iron Bank are flagged at the same 2.4% baseline rate as that of other traders, over an extended period of time.

**Test Statistic**

Our test statistic is the number of trades flagged by the SEC algorithm.


```{r, echo=FALSE}
ggplot(illegal_trade_sim) + 
  geom_histogram(aes(x=nflip), fill = "lavender", col = "burlywood", binwidth=1) + labs(title="Plot of the Number of Trades Flagged by the SEC Algorithm", subtitle="assuming that securities trades from the Iron Bank are flagged at the same 2.4% baseline rate\nas that of other traders, over an extended period of time")
```

**P-value**

The p-value we have derived from this data is 0.0021, or 210 out of 100,000. This p-value is quite small, meaning that only a small part of the distribution is 70 tagged trades or higher.

**Conclusion**

While this could mean nothing, there is a good chance that there is something here that the SEC officials may need to look into.

Thus, in this case, it seems that the null hypothesis that securities trades from the Iron Bank are flagged at the same 2.4% baseline rate as that of other traders, must be rejected. There is a high chance that suspicious activity may be occurring regarding the trades from the Iron Bank.

### Problem 2: Health Inspections

```{r, echo=FALSE, results = 'hide'}
health_inspection_sim = do(100000)*nflip(n=50, prob=0.03)
sum(health_inspection_sim >= 8) # p-val: 0.0001
14/100000
```

**Null Hypothesis**

The null hypothesis that we are testing is that on average, restaurants in the city are cited for health code violations at the same 3% baseline rate.

**Test Statistic**

Our test statistic is the number of health code violations given to Gourmet Bites.


```{r, echo=FALSE}
ggplot(health_inspection_sim) + 
  geom_histogram(aes(x=nflip), fill = "lavender", col = "burlywood", binwidth=1) + labs(title="Plot of the Number of Health Code Violations Given to Gourmet Bites", subtitle="assuming that on average, restaurants in the city are cited for health code violations at the same 3% baseline rate")
```

**P-value**

The p-value we have derived from this data is 0.00014, or 14 out of 100,000. This p-value is quite small, meaning that only a very small part of the distribution has 8 health code violations or higher.

**Conclusion**

As the p-value is very small, it seems that the null hypothesis that on average, restaurants in the city are cited for health code violations at the same 3% baseline rate, must be rejected. There is a high chance that the Health Department may have to take action as Gourmet Bites' rate of health code violations is significantly higher than the citywide average of 3%.

### Problem 3: LLM Watermarking

```{r, echo=FALSE, results = 'hide'}
# null distribution
brown_sent_txt <- readLines("brown_sentences.txt") 

brown_sent <- tibble(Sentences = brown_sent_txt)

calculate_observed_expected = function(sentence, freq_table) {
  
  # Ensure letter frequencies are normalized and sum to 1
  freq_table$Probability = freq_table$Probability / sum(freq_table$Probability)
  
  # Remove non-letters and convert to uppercase
  clean_sentence = gsub("[^A-Za-z]", "", sentence)
  clean_sentence = toupper(clean_sentence)
  
  # Count the occurrences of each letter in the sentence
  observed_counts = table(factor(strsplit(clean_sentence, "")[[1]], levels = freq_table$Letter))
  
  # Calculate expected counts
  total_letters = sum(observed_counts)
  expected_counts = total_letters * freq_table$Probability
  
  return(list(obs_counts = observed_counts, exp_counts = expected_counts))
}

brown_sent_null_distribution = numeric(nrow(brown_sent))

for (i in 1:nrow(brown_sent)) {
  
  brown_sent_residuals <- calculate_observed_expected(brown_sent$Sentences[i], freq_table = letter_freq)
  
  chi_squared_stat <- sum((brown_sent_residuals$obs_counts - brown_sent_residuals$exp_counts)^2 / brown_sent_residuals$exp_counts)
  
  brown_sent_null_distribution[i] <- chi_squared_stat
}

brown_sent$Chi_Squared <- brown_sent_null_distribution

```

```{r, echo=FALSE, results = 'hide'}
ggplot(brown_sent) +
  geom_histogram(aes(x = Chi_Squared), fill = "lavender", col = "burlywood", binwidth=1)
```

```{r, echo=FALSE, results = 'hide'}
test_sent <- c(
  "She opened the book and started to read the first chapter, eagerly anticipating what might come next.",
  "Despite the heavy rain, they decided to go for a long walk in the park, crossing the main avenue by the fountain in the center.",
  "The museum’s new exhibit features ancient artifacts from various civilizations around the world.",
  "He carefully examined the document, looking for any clues that might help solve the mystery.",
  "The students gathered in the auditorium to listen to the guest speaker’s inspiring lecture.",
  "Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland.",
  "The chef demonstrated how to prepare a delicious meal using only locally sourced ingredients, focusing mainly on some excellent dinner recipes from Spain.",
  "They watched the sunset from the hilltop, marveling at the beautiful array of colors in the sky.",
  "The committee reviewed the proposal and provided many points of useful feedback to improve the project’s effectiveness.",
  "Despite the challenges faced during the project, the team worked tirelessly to ensure its successful completion, resulting in a product that exceeded everyone’s expectations."
)
```

```{r, echo=FALSE, results = 'hide'}
test_sent <- tibble(Sentences = test_sent)

test_sent_distribution = numeric(nrow(test_sent))

for (i in 1:nrow(test_sent)) {
  
  test_sent_residuals <- calculate_observed_expected(test_sent$Sentences[i], freq_table = letter_freq)
  
  chi_squared_stat <- sum((test_sent_residuals$obs_counts - test_sent_residuals$exp_counts)^2 / test_sent_residuals$exp_counts)
  
  test_sent_distribution[i] <- chi_squared_stat
}

test_sent$Chi_Squared <- test_sent_distribution

test_sent$pval <- sapply(test_sent_distribution, function(ChiSquare){
  sum(brown_sent_null_distribution >= ChiSquare) / length(brown_sent_null_distribution)
})
```

```{r, echo=FALSE, results = 'hide'}
test_sent_pval <- test_sent %>%
  select(pval)

test_sent_pval <- test_sent_pval %>%
  mutate(pval = round(pval, 3))

test_sent_pval <- test_sent_pval %>%
  mutate(Sentence = c(1:10))

test_sent_pval <- test_sent_pval %>%
  select(Sentence, pval)

sentences_kable <- knitr::kable(test_sent_pval, col.names = c("Sentence Number", "P-value"), align = "l", caption = "English Sentence Letter Distribution")

sentences_kable <- sentences_kable %>% 
  kable_styling()

```

```{r, echo=FALSE}
sentences_kable
```

The sentence that is watermarked by asking the LLM to subtly adjust its frequency distribution over letters, is sentence 6. We can tell this as it is the smallest p-value, thus rejecting the null distribution.  
